### 2021-11-24 src/article.tex
Initial commit

%% The first command in your LaTeX source must be the \documentclass command.
%%
%% Options:
%% twocolumn : Two column layout.
%% hf: enable header and footer.
\documentclass[
    twocolumn,
% hf,
]{ceurart}

%%
%% One can fix some overfulls
\sloppy

%%
%% Minted listings support
%% Need pygment <http://pygments.org/> <http://pypi.python.org/pypi/Pygments>
\usepackage{minted}
%% auto break lines
\setminted{breaklines=true}

%%
%% end of the preamble, start of the body of the document source.
\begin{document}

%%
%% Rights management information.
%% CC-BY is default license.
    \copyrightyear{2021}
    \copyrightclause{Copyright for this paper by its authors.
    Use permitted under Creative Commons License Attribution 4.0
    International (CC BY 4.0).}

%%
%% This command is for the conference information
    \conference{Proceedings of the 13th Majorov International Conference on Software Engineering
    and Computer Systems,

        December 1--3, 2021, Online \& Saint Petersburg, Russia}

%%
%% The "title" command
    \title{Sparse Representation Approaches for First Stage Retrieval: A Review}

%%
%% The "author" command and its associated commands are used to define
%% the authors and their affiliations.
    \author[1]{Vyacheslav YU. Dobrynin}[%
        email=vidobrynin@itmo.ru
    ]

    \author[1]{Alexey V. Platonov}[%
        email=avplatonov@itmo.ru
    ]

    \address[1]{ITMO University, Kronverksky Pr. 49, bldg. A, Saint-Petersburg, 197101,
        Russian Federation}

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
    \begin{abstract}
        Modern information retrieval models have a two-stage architecture. This is due to the
        requirements - to search for the most relevant documents in a short time. For a long time,
        the attention of researchers was focused on the second stage, the stage of ranking a small
        corpus of documents obtained as a result of the work of the first stage. Thanks to neural
        networks, significant results have been achieved.

        Recently, researchers have begun to pay attention to the first stage, since it is important
        to form as many documents as possible corresponding to a search query for subsequent
        ranking. Due to the fact that at this stage it is necessary to work with millions or
        billions of documents, work efficiency is very important. An inverted index is very well
        suited for the task of quickly searching a huge dataset.

        In this paper, we look at modern sparse representation approaches for constructing an
        inverted index.
    \end{abstract}

%%
%% Keywords. The author(s) should pick words that accurately describe
%% the work being presented. Separate the keywords with commas.
    \begin{keywords}
        Information retrieval \sep
        inverted index \sep
        sparse representation \sep
        neural network
    \end{keywords}

%%
%% This command processes the author and affiliation and title
%% information and builds the first part of the formatted document.
    \maketitle

    \section{Introduction}

    Information retrieval predates the web. Its evolution was stimulated by various problems
    associated with providing search and access to information sources \cite{manning2008introduction}.
    Applications of this discipline are used to access information in databases of various
    organizations.

    But one of the main engines of progress is web search. Now, for most people, this is the main
    place to find information.

    The requirements for search engines are growing every year, the number of documents is growing,
    and the competition in the market is also high. Therefore, it is necessary to improve the
    quality of the search.

    For a long time, researchers have focused on the ranking stage, but recently, more and more
    papers have appeared that investigate ways to retrieve more relevant documents
    in the first stage.

    Significant results in natural language processing have been achieved through
    the use of embeddings, numerical representations of words or other linguistic entities.
    At the ranking stage, dense embeddings are successfully applied; neural networks are used
    to form them. But this approach is computationally ineffective due to the large number
    of documents that need to be worked with at the first stage. Therefore, many researchers pay
    their attention to sparse embeddings.

    In this paper, we will consider some of the currently existing methods for working with sparse
    representations.

    \section{Problem statement}

    \subsection{Modern information retrieval system architecture}

    To combine efficiency and effectiveness when searching for documents in a large corpus, modern
    search engines use multi-stage architectures as shown in the Figure \ref{fig:architecture}.
    \begin{figure}[h]
        \centering
        \includegraphics[width=\linewidth]{Architecture.png}
        \caption{The multi-stage architecture of modern information retrieval systems.}
        \label{fig:architecture}
    \end{figure}

    In the first stage, during a search in a large body of documents, the initial set of candidate
    documents is retrieved by using a cheaper ranking model with an inverse index. During the
    second stage, more accurate but slow models are used to reduce the set of documents obtained
    from the first stage and leave the most suitable. This approach is widely used in both academic
    and industrial fields. Moreover, it achieved state-of-the-art results.

    \subsection{Problem formalization} \label{problem-formalization}

    The task at the first stage of extraction is to train the model on the dataset, which will
    evaluate the relevance pairs consisting of a search query and a document. The model will give
    high scores to pairs that have a high relevance and low scores to pairs that do not. Thus,
    for any query-document pair (\emph{q}, \emph{d}), model \emph{s}(\emph{q},\emph{d}) calculates
    the degree of relevance between the query and the document. The scoring function is usually
    represented as follows:
    \begin{equation}
        s(q,d) = f(\phi(q), \psi(d)),
    \end{equation}
    where \emph{q} \in \emph{X} and \emph{d} \in \emph{Y} are the query and document. $\phi$ and
    $\psi$ are representation functions for map tokens in \emph{X} and \emph{Y} to their
    embeddings $\phi(q)$ and $\psi(d)$.
    To build a model, you need to fulfill the following requirements:
    \begin{itemize}
        \item The document representation function $\psi$ should be independent of the query,
        this allows it to be precomputed and indexed.
        \item The query representation function $\phi$ and scoring function \emph{f} should be as
        efficient as possible, since they are calculated online \cite{SOTA}.
    \end{itemize}

    \subsection{Sparse representation}

    Sparse retrieval methods usually represent each document and each query with sparse vectors,
    where only a small number of dimensions are active. The sparse representation has attracted
    great attention as it connects to the nature of human memories and shows better interpretability.

    Besides, sparse representations can be easily integrated into existing inverted indexing
    engines for efficient retrieval.

    Without loss of generality, sparse retrieval methods can be categorized into two classes.
    One is to encode queries and documents still in the symbolic space but employ neural models
    to improve term weighting schemes, namely neural weighting schemes. The other is to directly
    learn sparse representations, i.e., $\phi(q)$ and $\psi(d)$, in latent space for queries
    and documents with neural networks, which we call sparse representation learning \cite{SOTA}.

    \section{Models}

    \subsection{SNRM - standalone neural ranking model}

    \begin{figure}[h]
        \centering
        \def\svgwidth{\columnwidth}
        \input{SNRM.pdf_tex}
        \caption{Learning Latent Sparse Representation}
    \end{figure}

    The approach is based on the statement - if you build a neural learning model that meets the
    formal requirements in section \ref{problem-formalization}, then you can build an inverted
    index from the learned representation $\psi(d)$ and receive documents in response to a query
    without a first stage retrieval model.

    Neural network architecture is designed so that parameters can be shared between components
    $\phi(q)$ and $\psi(d)$, which as a result have one semantic space.

    Also, the representations of documents and the query have different levels of sparsity,
    basically for the purpose of greater efficiency, and this is also due to the fact that
    the query is usually much smaller than the document and it is logical to present it as
    representation with fewer elements.

    Therefore, a model is needed in which the representation sparsity is a function
    of the input length. Architecture based on Ngram representation learning. For each continuous
    n words in a given document or query, sparse representations are trained, which are then
    aggregated via the average pooling. Thus, the document similarity is represented as follows:
    \begin{equation}
        \phi_D(q)=\frac{1}{|d| - n + 1} \sum_{i=1}^{|d| - n + 1} \phi_{ngram}(w_i, w_{i+1}, \cdots, w_{i+n-1})
    \end{equation}
    where $w_1, w_2, \cdots, w_{|d|}$ - terms in document, $\phi_{ngram}$ - learns a sparse representation
    for the given ngram. The same function is used for query representation.
    Ngram representation function is a fully-connected feed-forward network that uses a sliding
    window to read the input words and encodes their ngrams. Then embeddings are collected
    for each term \cite{SNRM}.

    \paragraph{Loss Function}
    The loss function for the $i^{th}$ training instance has the following form:
    \begin{equation}
        \mathcal{L}(q_i,d_{i1},d_{i2},y_i) + \lambda\;L_1(\phi_Q(q_i)||\phi_D(d_{i1})||\phi_D(d_{i2}))
    \end{equation}
    where $\mathcal{L}$ - hinge loss \cite{IRNLP}, $q_i$ - query, $d_{i1}$ and $d_{i2}$ - two
    document candidates, $y_i \in \{-1, 1\}$ - label that shows the relevance between the document
    and the query, $\lambda$ - the hyper-parameter to control the sparsity of the learned
    representations, $L_1$ is norm (i.e.
    \begin{math}
        L_1(\vec{\upsilon}) = \sum_{i=1}^{|\vec{\upsilon}|}|\vec{\upsilon}_i|
    \end{math}
    ) and $||$ is concatenation sign.\\\par

    However, SNRM has disadvantage, namely, it loses the interpretability of the original terms,
    which can be significant for industrial systems.

    \subsection{SparTerm}

    \begin{figure}[h]
        \centering
        \def\svgwidth{\columnwidth}
        \input{SparTerm.pdf_tex}
        \caption{General architecture of SparTerm model}
    \end{figure}
    The SparTerm model includes an importance predictor $\mathnormal{F}$ and a gating controller
    $\mathnormal{G}$, it maps the original textual passage \emph{p} to a sparse representation
    $p'$ in the vocabulary. This process can be represented as follows:

    \begin{equation}
        \label{eqn:sparterm:mapping-process}
        p' = \mathnormal{F}(p) \odot \mathnormal{G}(p)
    \end{equation}

    $\mathnormal{F}$ generates a dense vector representing the semantic importance of each item
    in the vocabulary. $\mathnormal{G}$ generates a binary gating vector to control which terms
    to appear in the final sparse representation.

    The importance $w_{ij}$ of the token \emph{j} (vocabulary) for token \emph{i} (of the input
    sequence) presented as follows:
    \begin{equation}
        w_{ij}=transform(h_i)^T E_j + b_j \quad j \in {1, ..., |V|}
    \end{equation}
    where \emph{transform(.)} denotes a linear transformation with GeLU
    (Gaussian Error Linear Unit \cite{hendrycks2016gelu}) activation and layer normalization,
    $E_j$ is the BERT input embeddings for token \emph{j} and $b_j$ is the bias term.
    The final representation is then received by the summation of all token-wise importance
    distributions, after applying ReLU to ensure the positivity of term weights:
    \begin{equation}
        \label{eqn:sparterm:importance-predictor}
        w_j=\sum_{i=0}^L ReLU(w_{ij})
    \end{equation}
    where L is the input sequence length.

    \paragraph{Training the importance predictor.}
    There is a set of N training instances $R=\{(q_1, p_{1+}, p_{1-}), ..., (q_N, p_N+, p_N-)\}$,
    each of which contains a query $q_i$, a positive passage $p_{i+}$ and a negative one $p_i-$,
    such that the first passage is more relevant the query than the second one. As a result,
    the loss function has the following form:
    \begin{equation}
        L_{rank}(q_i,p_{i+},p_{i-}) = -log \frac
        {e^{sim(q_i',p_{i+}')}}
        {e^{sim(q_i',p_{i+}')} + e^{sim(q_i',p_{i-}')}}
    \end{equation}
    where $q_i',p_{i+}',p_{i-}'$ is the sparse representation of $q_i,p_{i+},p_{i-}$
    received by Equation (\ref{eqn:sparterm:mapping-process}), \emph{sim} represents a function
    to measure the similarity (e.g. dot product).

    \paragraph{Training the gating controller.}
    Consider a passage-query/summary parallel corpus \emph{C}, which contains passage \emph{p}, the
    cor- responding target text \emph{t}, and the binary bag-of-words vector of \emph{t} (\emph{T}).
    In order to maximize probability values of all terms in vocabulary, binary cross-entropy loss
    is used, which has the following form:
    \begin{equation}
        L_{exp}=-\lambda_1\sum_{j \in \{m|T_m=0\}}log(1-G_j)-\lambda_2\sum_{k \in \{m|T_m=1\}}logG_k
    \end{equation}
    where $\lambda_1$ and $\lambda_2$ are tunable hyper-parameters, first is for terms expected not to be
    expanded, second one is for terms that appear in the target text, and \emph{G} is is the dense
    gating probability distribution for \emph{p} \cite{SparTerm}.\\\par
    Thus, SparTerm provides a function that is used to map the frequency-based and bag-of-words
    representation to a sparse term importance distribution in the whole vocabulary. This model
    also has limitation: since the gating mechanism is learned in advance, this prevents
    the model from learn the optimal sparsification strategy for the ranking task.

    \subsection{SPLADE}

    This approach is based on the SparTerm model, it allows to improve it and get rid
    of the limitation described above.

    \paragraph{Architecture}
    The importance estimation formula Equation (\ref{eqn:sparterm:importance-predictor})
    has been modified by adding a log-saturation effect, in order to prevent the dominance
    of some terms to dominate and naturally ensures sparsity in representations:
    \begin{equation}
        w_j=\sum_{i \in t}log(1+ReLU(w_{ij}))
    \end{equation}

    \paragraph{Loss function}
    The loss function has the following form:
    \begin{equation}
        L_{rank-IBN}=-log
        \frac{
            e^{s(q_i,d_i^+)}
        }{
            e^{s(q_i,d_i^+)} +
            e^{s(q_i,d_i^-)} +
            \sum_j e^{s(q_i,d_{i,j}^-)}
        }
    \end{equation}
    where IBN is in-batch negatives sampling strategy, which is effective for learning first-stage
    rankers \cite{IBN1,IBN2,IBN3}, $q_i$ is query, $d_i^+$ and $d_i^-$ are positive and negative
    documents respectively, ${d_{i,j}^-}$ is set of positive documents from other queries,
    but negative for the current one \cite{SPLADE}.\\\par
    As a result, SPLADE is simple and effective model in information retrieval tasks.

    \subsection{RepBERT}

    For comparison, consider the dense retrieval method called a RepBERT based
    on the popular BERT model.

    \paragraph{Architecture}
    As in BERT\cite{BERT}, the input text is wrapped in [CLS] and [SEP] service marks
    during tokenization:

    \begin{equation}
        Input(text) = [CLS]\;Tokenize(text)\;[SEP]
    \end{equation}

    RepBERT uses an encoder, that finds the averaged value of contextualized vectors for each token
    obtained by passing it to BERT:

    \begin{multline}
        Embed(text)=Encoder(text)=\\
        Average(BERT(Input(text)))
    \end{multline}

    After obtaining the embeddings of queries and documents, RepBERT shares their weights
    to receive relevance scores:

    \begin{equation}
        Rel(query,doc) = Embed(query)^T \cdot Embed(doc)
    \end{equation}

    \paragraph{Training}
    During training, it is necessary to make a embeddings inner product for the relevant
    query-document pairs larger than for the irrelevant ones. The model uses the
    MultiLabelMarginLoss as loss function:
    \begin{multline}
        \mathnormal{L}(q,d_1^+,...,d_m^+,d_{m+1}^-,...,d_n^-) = \\
        \frac{1}{n} \cdot \sum_{1 \leq i \leq m,m < j \leq n}
        max(0, 1 - (Rel(q,d_i^+) - Rel(q,d_j^-)))
    \end{multline}
    where $(q,d_1^+,...,d_m^+,d_{m+1}^-,...,d_n^-)$ is one instance of the input training batch,
    \emph{q} is query, \emph{m} is quantity of relevant documents and \emph{n - m} is quantity
    of irrelevant documents\cite{RepBERT}.\\\par
    The model has a simple architecture in order to improve work efficiency.

% \subsection{UHD-BERT}
% A novel sparse retrieval method empowered by extremely high dimensionality and controllable
%     sparsity. To fully utilize the abundant capacity of representation scheme, proposed bucketed
%     learning, which is a method of learning from the same or different layers of BERT
%     for different aspects or abstraction levels of information.
%     Also introduce and justify choice of WTA model as sparsification model by proving the
%     benefits of reliable representation generation and the ability to control the sparsity
%     at inference time. Show that our model outperforms previous sparse models significantly
%     and delivers competitive performance compared to dense models especially for recall,
%     which is an important measure for a first-stage ranker.

    \section{Evaluation}

    Let's bring the retrieval accuracy for models SparTerm, SPLADE, RepBERT into a pivot table.
    The SNRM model is older and more basic, so it will not be considered.

    To assess the accuracy, the MS MARCO dataset\cite{MS-MARCO} is used.

    The comparison results are shown in the Table \ref{tab:evaluation}.
    Mean Reciprocal Rank (MMR@10) is used as a metric for assessing accuracy.

    \begin{table}[h]
        \caption{The retrieval accuracy (MRR@10) of different approaches. Dataset: MS MARCO.}
        \label{tab:evaluation}
        \begin{tabular}{p{0.4\columnwidth}|p{0.4\columnwidth}}
            \toprule
            Model                   & MRR@10 \\
            \midrule
            SparTerm\cite{SparTerm} & 0.290  \\
            SPLADE\cite{SPLADE}     & 0.322  \\
            RepBERT\cite{RepBERT}   & 0.304  \\
            \bottomrule
        \end{tabular}
    \end{table}

    \section{Conclusion}

    The purpose of this review is to summarize the current research status on sparse representation
    approaches and gain some insights for future development.
    It includes a brief review of existing state-of-the-art sparse retrieval methods, as well as a
    dense retrieval method for an ad-hoc retrieval task. A brief description of the approaches
    is given, illustrations are given, and the necessary mathematical equations are outlined.

    At the end, a comparison was made of the approaches considered, from which it can be seen that
    the SPLADE method has the highest accuracy, followed by the compressed RepBERT, but this
    approach are less effective due to the fact that it is based on dense representations.

    The information provided will be used for further research. Our focus will be more on sparse
    representations for the first stage of information retrieval. In further works, it is assumed
    that we have our own stand for researching deep learning models in order to form an index
    in an information retrieval system.

%%
%% Define the bibliography file to be used
    \bibliography{article}

%%
%% If your work has an appendix, this is the place to put it.
    \appendix

\end{document}

%%
%% End of file

### 2021-11-28 src/article.tex
Update section [Problem formalization]
Each sentence on new line
Update section [Problem formalization]. Update and refactoring bibliography.

@@
     \begin{abstract}
-        Modern information retrieval models have a two-stage architecture. This is due to the
-        requirements - to search for the most relevant documents in a short time. For a long time,
-        the attention of researchers was focused on the second stage, the stage of ranking a small
-        corpus of documents obtained as a result of the work of the first stage. Thanks to neural
-        networks, significant results have been achieved.
+        Modern information retrieval models have a two-stage architecture.
+        This is due to the requirements - to search for the most relevant documents in a short time.
+        For a long time, the attention of researchers was focused on the second stage, the stage
+        of ranking a small corpus of documents obtained as a result of the work of the first stage.
+        Thanks to neural networks, significant results have been achieved.
 
@@
         to form as many documents as possible corresponding to a search query for subsequent
-        ranking. Due to the fact that at this stage it is necessary to work with millions or
-        billions of documents, work efficiency is very important. An inverted index is very well
-        suited for the task of quickly searching a huge dataset.
+        ranking.
+        Due to the fact that at this stage it is necessary to work with millions or billions of
+        documents, work efficiency is very important.
+        An inverted index is very well suited for the task of quickly searching a huge dataset.
 
@@
 
-    Information retrieval predates the web. Its evolution was stimulated by various problems
-    associated with providing search and access to information sources \cite{manning2008introduction}.
+    Information retrieval predates the web.
+    Its evolution was stimulated by various problems associated with providing search and access
+    to information sources \cite{manning2008introduction}.
     Applications of this discipline are used to access information in databases of various
@@
 
-    But one of the main engines of progress is web search. Now, for most people, this is the main
-    place to find information.
+    But one of the main engines of progress is web search.
+    Now, for most people, this is the main place to find information.
 
     The requirements for search engines are growing every year, the number of documents is growing,
-    and the competition in the market is also high. Therefore, it is necessary to improve the
-    quality of the search.
+    and the competition in the market is also high.
+    Therefore, it is necessary to improve the quality of the search.
 
@@
     the use of embeddings, numerical representations of words or other linguistic entities.
-    At the ranking stage, dense embeddings are successfully applied; neural networks are used
-    to form them. But this approach is computationally ineffective due to the large number
-    of documents that need to be worked with at the first stage. Therefore, many researchers pay
-    their attention to sparse embeddings.
+    At the ranking stage, dense embeddings are successfully applied;
+    neural networks are used to form them.
+    But this approach is computationally ineffective due to the large number of documents that
+    need to be worked with at the first stage.
+    Therefore, many researchers pay their attention to sparse embeddings.
 
@@
     In the first stage, during a search in a large body of documents, the initial set of candidate
-    documents is retrieved by using a cheaper ranking model with an inverse index. During the
-    second stage, more accurate but slow models are used to reduce the set of documents obtained
-    from the first stage and leave the most suitable. This approach is widely used in both academic
-    and industrial fields. Moreover, it achieved state-of-the-art results.
+    documents is retrieved by using a cheaper ranking model with an inverse index.
+    During the second stage, more accurate but slow models are used to reduce the set of
+    documents obtained from the first stage and leave the most suitable.
+    This approach is widely used in both academic and industrial fields.
+    Moreover, it achieved state-of-the-art results.
 
@@
 
+    Ideally, the goal of the first stage of a search is to retrieve all relevant documents from the
+    collection.
+    In this regard, when building a model, it is necessary to strive to ensure that it can return
+    the most complete set of relevant documents in a short period of time.
+
+    For a long time, term-based models were used for the first stage of search, which showed high
+    efficiency due to simple logic and the use of an inverse index
+    \cite{termBasedModelEfficiency1, termBasedModelEfficiency2}.
+
     The task at the first stage of extraction is to train the model on the dataset, which will
-    evaluate the relevance pairs consisting of a search query and a document. The model will give
-    high scores to pairs that have a high relevance and low scores to pairs that do not. Thus,
-    for any query-document pair (\emph{q}, \emph{d}), model \emph{s}(\emph{q},\emph{d}) calculates
-    the degree of relevance between the query and the document. The scoring function is usually
-    represented as follows:
+    evaluate the relevance pairs consisting of a search query and a document.
+    The model will give high scores to pairs that have a high relevance and low scores to
+    pairs that do not.
+    Thus, for any query-document pair (\emph{q}, \emph{d}), model \emph{s}(\emph{q},\emph{d})
+    calculates the degree of relevance between the query and the document.
+    The scoring function is usually represented as follows:
     \begin{equation}
@@
     Sparse retrieval methods usually represent each document and each query with sparse vectors,
-    where only a small number of dimensions are active. The sparse representation has attracted
-    great attention as it connects to the nature of human memories and shows better interpretability.
+    where only a small number of dimensions are active.
+    The sparse representation has attracted great attention as it connects to the nature of human
+    memories and shows better interpretability.
 
@@
     One is to encode queries and documents still in the symbolic space but employ neural models
-    to improve term weighting schemes, namely neural weighting schemes. The other is to directly
-    learn sparse representations, i.e., $\phi(q)$ and $\psi(d)$, in latent space for queries
-    and documents with neural networks, which we call sparse representation learning \cite{SOTA}.
+    to improve term weighting schemes, namely neural weighting schemes.
+    The other is to directly learn sparse representations, i.e., $\phi(q)$ and $\psi(d)$, in
+    latent space for queries and documents with neural networks, which we call sparse
+    representation learning \cite{SOTA}.
 
@@
     Therefore, a model is needed in which the representation sparsity is a function
-    of the input length. Architecture based on Ngram representation learning. For each continuous
-    n words in a given document or query, sparse representations are trained, which are then
-    aggregated via the average pooling. Thus, the document similarity is represented as follows:
+    of the input length.
+    Architecture based on Ngram representation learning.
+    For each continuous n words in a given document or query, sparse representations are trained,
+    which are then aggregated via the average pooling.
+    Thus, the document similarity is represented as follows:
     \begin{equation}
@@
     where $w_1, w_2, \cdots, w_{|d|}$ - terms in document, $\phi_{ngram}$ - learns a sparse representation
-    for the given ngram. The same function is used for query representation.
+    for the given ngram.
+    The same function is used for query representation.
     Ngram representation function is a fully-connected feed-forward network that uses a sliding
-    window to read the input words and encodes their ngrams. Then embeddings are collected
-    for each term \cite{SNRM}.
+    window to read the input words and encodes their ngrams.
+    Then embeddings are collected for each term \cite{SNRM}.
 
@@
     $\mathnormal{G}$, it maps the original textual passage \emph{p} to a sparse representation
-    $p'$ in the vocabulary. This process can be represented as follows:
+    $p'$ in the vocabulary.
+    This process can be represented as follows:
 
@@
     each of which contains a query $q_i$, a positive passage $p_{i+}$ and a negative one $p_i-$,
-    such that the first passage is more relevant the query than the second one. As a result,
-    the loss function has the following form:
+    such that the first passage is more relevant the query than the second one.
+    As a result, the loss function has the following form:
     \begin{equation}
@@
     Thus, SparTerm provides a function that is used to map the frequency-based and bag-of-words
-    representation to a sparse term importance distribution in the whole vocabulary. This model
-    also has limitation: since the gating mechanism is learned in advance, this prevents
+    representation to a sparse term importance distribution in the whole vocabulary.
+    This model also has limitation: since the gating mechanism is learned in advance, this prevents
     the model from learn the optimal sparsification strategy for the ranking task.
@@
     During training, it is necessary to make a embeddings inner product for the relevant
-    query-document pairs larger than for the irrelevant ones. The model uses the
-    MultiLabelMarginLoss as loss function:
+    query-document pairs larger than for the irrelevant ones.
+    The model uses the MultiLabelMarginLoss as loss function:
     \begin{multline}
@@
     It includes a brief review of existing state-of-the-art sparse retrieval methods, as well as a
-    dense retrieval method for an ad-hoc retrieval task. A brief description of the approaches
-    is given, illustrations are given, and the necessary mathematical equations are outlined.
+    dense retrieval method for an ad-hoc retrieval task.
+    A brief description of the approaches is given, illustrations are given, and the necessary
+    mathematical equations are outlined.
 
@@
 
-    The information provided will be used for further research. Our focus will be more on sparse
-    representations for the first stage of information retrieval. In further works, it is assumed
-    that we have our own stand for researching deep learning models in order to form an index
-    in an information retrieval system.
+    The information provided will be used for further research.
+    Our focus will be more on sparse representations for the first stage of information retrieval.
+    In further works, it is assumed that we have our own stand for researching deep learning
+    models in order to form an index in an information retrieval system.
 

### 2021-11-29 src/article.tex
Update section [Problem formalization]

@@
     \cite{termBasedModelEfficiency1, termBasedModelEfficiency2}.
+    However, this approach has drawbacks in the form of vocabulary mismatch
+    \cite{vocabularyMismatch1, vocabularyMismatch2} and lack of consideration of the order of
+    terms in documents \cite{ignoringTermOrdering}.
+    Thus, term-based models can lose some of the relevant documents at the very beginning.
+    To get rid of these limitations, various solutions have been proposed, such as vocabularies
+    of synonyms or query \cite{queryExp1,queryExp2,queryExp3,queryExp4} and document
+    \cite{docExp1, docExp2, docExp3} expansion.
+    However, progress has been limited due to the fact that these approaches still remain in
+    the term-based paradigm and inherit its limitations.
+
+    To solve this problem, you can train a neural network that will encode the query and the
+    document so that they are in the same semantic space. Most often such spaces are dense
+    representations, but they are algorithmically complex because it is necessary to iterate over
+    the vectors of documents during the search. In this regard, the researchers decided to
+    represent documents and queries in vector spaces of higher dimension (sparse representations)
+    with their further application on the inverse index.
 
@@
     \end{itemize}
+
+    An example is the BM25 \cite{bm25} algorithm which is a retrieval function based on a bag of
+    words.
+    The scoring function uses TF-IDF weights and term matching at its core in order to
+    determine the similarity of a document and a query, with its help at the first stage of
+    the search, candidate documents are retrieved for subsequent re-ranking.
 

### 2021-12-03 src/article.tex
Update section [Sparse representation]

@@
     memories and shows better interpretability.
+    Since usually the term is replaced with a numeric value and has a one-to-one correspondence
+    in the vocabulary of words.
+    This feature makes it easy to use sparse representations on an inverted index, in this case
+    each value of the term map to the document in which it occurs.
 
-    Besides, sparse representations can be easily integrated into existing inverted indexing
-    engines for efficient retrieval.
+    During the search, the query is broken down into terms in the same way as documents, then
+    they are used as keys in the inverted index to find the corresponding documents.
+    The more query terms are found in a document, the more suitable it is for re-ranking.
 

### 2021-12-03 src/article.tex
Rename section [Evaluation -> Models comparison]
Add bibliography about invertedIndexTechnique

@@
     For a long time, term-based models were used for the first stage of search, which showed high
-    efficiency due to simple logic and the use of an inverse index
-    \cite{termBasedModelEfficiency1, termBasedModelEfficiency2}.
+    efficiency \cite{termBasedModelEfficiency1, termBasedModelEfficiency2} due to simple logic
+    and the use of an inverse index.
     However, this approach has drawbacks in the form of vocabulary mismatch
@@
     The more query terms are found in a document, the more suitable it is for re-ranking.
+    For details on the inverted index technique, see
+    \cite{invertedIndexTechnique1, invertedIndexTechnique2}
 
@@
 
-    \section{Evaluation}
+    \section{Models comparison}
 
@@
     The comparison results are shown in the Table \ref{tab:evaluation}.
-    Mean Reciprocal Rank (MMR@10) is used as a metric for assessing accuracy.
+    Mean Reciprocal Rank (MRR@10) is used as a metric for assessing accuracy.
 

### 2021-12-03 src/article.tex
Add bibliography about embeddings

@@
     Significant results in natural language processing have been achieved through
-    the use of embeddings, numerical representations of words or other linguistic entities.
+    the use of embeddings \cite{embedding1, embedding2, embedding3}, numerical representations of
+    words or other linguistic entities.
     At the ranking stage, dense embeddings are successfully applied;

### 2021-12-06 src/article.tex
Figure for SPLADE

@@
     \paragraph{Architecture}
+    \begin{figure}[h]
+        \centering
+        \def\svgwidth{\columnwidth}
+        \input{SPLADE.pdf_tex}
+        \caption{Final representation calculation with log saturation}
+        \label{fig:SPLADE}
+    \end{figure}
     The importance estimation formula Equation (\ref{eqn:sparterm:importance-predictor})
-    has been modified by adding a log-saturation effect, in order to prevent the dominance
-    of some terms to dominate and naturally ensures sparsity in representations:
+    has been modified by adding a log-saturation effect (Figure \ref{fig:SPLADE}), in order to
+    prevent the dominance of some terms to dominate and naturally ensures sparsity in
+    representations:
     \begin{equation}

### 2021-12-11 src/article.tex
Figure for RepBERT

@@
     \paragraph{Architecture}
+    \begin{figure}[h]
+        \centering
+        \includegraphics[scale=0.2]{RepBERT.png}
+        \caption{RepBERT architecture}
+        \label{fig:RepBERT}
+    \end{figure}
+    The process of receiving embeddings from document or query sentences is shown
+    in the Figure \ref{fig:RepBERT}.
+
     As in BERT\cite{BERT}, the input text is wrapped in [CLS] and [SEP] service marks

### 2021-12-14 src/article.tex
Fix typo.

@@
     Consider a passage-query/summary parallel corpus \emph{C}, which contains passage \emph{p}, the
-    cor- responding target text \emph{t}, and the binary bag-of-words vector of \emph{t} (\emph{T}).
+    corresponding target text \emph{t}, and the binary bag-of-words vector of \emph{t} (\emph{T}).
     In order to maximize probability values of all terms in vocabulary, binary cross-entropy loss

### 2021-12-19 src/article.tex
Add MultiLabelMarginLoss bibl
Update RepBERT score function.
Update SparTerm L_rank function.

@@
     \paragraph{Training the importance predictor.}
-    There is a set of N training instances $R=\{(q_1, p_{1+}, p_{1-}), ..., (q_N, p_N+, p_N-)\}$,
-    each of which contains a query $q_i$, a positive passage $p_{i+}$ and a negative one $p_i-$,
-    such that the first passage is more relevant the query than the second one.
+    There is a set of N training instances $R=\{(q_1, d_{1+}, d_{1-}), ..., (q_N, d_N+, d_N-)\}$,
+    each of which contains a query $q_i$, a positive document (passage) $d_{i+}$ and a negative
+    one $d_i-$, such that the first document is more relevant the query than the second one.
     As a result, the loss function has the following form:
     \begin{equation}
-        L_{rank}(q_i,p_{i+},p_{i-}) = -log \frac
-        {e^{sim(q_i',p_{i+}')}}
-        {e^{sim(q_i',p_{i+}')} + e^{sim(q_i',p_{i-}')}}
+        L_{rank} = -log \frac
+        {e^{s(q_i,d_{i+})}}
+        {e^{s(q_i,d_{i+})} + e^{s(q_i,d_{i-})}}
     \end{equation}
-    where $q_i',p_{i+}',p_{i-}'$ is the sparse representation of $q_i,p_{i+},p_{i-}$
-    received by Equation (\ref{eqn:sparterm:mapping-process}), \emph{sim} represents a function
-    to measure the similarity (e.g. dot product).
+    where $s(q,d)$ is denote the ranking score obtained via dot product between $d$ and $q$
+    representations from Equation (\ref{eqn:sparterm:mapping-process}).
 
@@
     \begin{equation}
-        Rel(query,doc) = Embed(query)^T \cdot Embed(doc)
+        s(q,d) = Embed(q)^T \cdot Embed(d)
     \end{equation}
+    where $Embed(q)^T$ and $Embed(d)$ are $\phi(q)$ and $\psi(d)$ respectively.
 
@@
     query-document pairs larger than for the irrelevant ones.
-    The model uses the MultiLabelMarginLoss as loss function:
+    The model uses the MultiLabelMarginLoss \cite{multiLabelMarginLoss} as loss function:
     \begin{multline}

### 2021-12-19 src/article.tex
Update SPLADE loss function.

@@
     \begin{equation}
+        \label{eqn:sparterm:loss}
         L_{rank} = -log \frac
@@
     \begin{equation}
+        \label{eqn:splade:importance}
         w_j=\sum_{i \in t}log(1+ReLU(w_{ij}))
@@
     \paragraph{Loss function}
-    The loss function has the following form:
+    The function presented in the Equation (\ref{eqn:sparterm:loss}) is supplemented with negative
+    sampling as shown below:
     \begin{equation}
@@
     documents respectively, ${d_{i,j}^-}$ is set of positive documents from other queries,
-    but negative for the current one \cite{SPLADE}.\\\par
+    but negative for the current one \cite{SPLADE}.
+
+    This approach does not use binary gating as in SparTerm, and instead learn model (Equation
+    \ref{eqn:splade:importance}) jointly optimizing ranking and regularization losses:
+    \begin{equation}
+        L = L_{rank-IBN} + \lambda_q L^q_{reg} + \lambda_d L^d_{reg}
+    \end{equation}
+    where $L_{reg}$ is a sparse regularisation ($L_1$ or $L_{FLOPS}$), $\lambda_q$ and
+    $\lambda_d$ are tunable hyper-parameters to control the sparsity of the query and the
+    document. \newline
+
     As a result, SPLADE is simple and effective model in information retrieval tasks.

### 2021-12-26 src/article.tex
Update conclusion.

@@
     At the end, a comparison was made of the approaches considered, from which it can be seen that
-    the SPLADE method has the highest accuracy, followed by the compressed RepBERT, but this
-    approach are less effective due to the fact that it is based on dense representations.
+    the SPLADE method has the highest accuracy, followed by the RepBERT, but this
+    approach are less effective due to the fact that it is based on dense representations the
+    embeddings for which are stored in a relatively slow vector database.
+    However, SPLADE is slower to learn than SparTerm due to the use of negative sampling.
 
-    The information provided will be used for further research.
-    Our focus will be more on sparse representations for the first stage of information retrieval.
-    In further works, it is assumed that we have our own stand for researching deep learning
-    models in order to form an index in an information retrieval system.
+    In further works, it is assumed that we have our own stand.
+    The obtained stand is planned to be used for further research, the purpose of which is to
+    develop an algorithm based on sparse representations with a search accuracy similar to
+    current state-of-the-art, but requiring less computational resources and memory.
 

### 2022-01-03 src/article.tex
Update SNRM and SparTerm score function info.

@@
     \begin{equation}
-        \phi_D(q)=\frac{1}{|d| - n + 1} \sum_{i=1}^{|d| - n + 1} \phi_{ngram}(w_i, w_{i+1}, \cdots, w_{i+n-1})
+        \psi(d)=
+        \frac{1}{|d| - n + 1}
+        \sum_{i=1}^{|d| - n + 1}
+        \psi_{ngram}(w_i, w_{i+1}, \cdots, w_{i+n-1})
     \end{equation}
-    where $w_1, w_2, \cdots, w_{|d|}$ - terms in document, $\phi_{ngram}$ - learns a sparse representation
+    where $w_1, w_2, \cdots, w_{|d|}$ - terms in document, $\psi_{ngram}$ - learns a sparse representation
     for the given ngram.
+
     The same function is used for query representation.
+
+    The matching function for score is defined as the dot product of the document and query
+    representations.
+
     Ngram representation function is a fully-connected feed-forward network that uses a sliding
@@
     \begin{equation}
-        \mathcal{L}(q_i,d_{i1},d_{i2},y_i) + \lambda\;L_1(\phi_Q(q_i)||\phi_D(d_{i1})||\phi_D(d_{i2}))
+        \mathcal{L}(q_i,d_{i1},d_{i2},y_i) + \lambda\;L_1(\phi(q_i)||\psi(d_{i1})||\psi(d_{i2}))
     \end{equation}
@@
     expanded, second one is for terms that appear in the target text, and \emph{G} is is the dense
-    gating probability distribution for \emph{p} \cite{SparTerm}.\\\par
+    gating probability distribution for \emph{p} \cite{SparTerm}.
+
+    \paragraph{Joint training.}
+    The importance predictor and gating controller are trained jointly because the supervisory
+    ranking signal can also be leveraged to guide the training of the gating controller:
+    \begin{equation}
+        L = L_{rank} + L_{exp}
+    \end{equation}\\
+
     Thus, SparTerm provides a function that is used to map the frequency-based and bag-of-words

### 2022-01-04 src/article.tex
Update loss function paragraphs.

@@
 
-    \paragraph{Loss Function}
+    \paragraph{Loss function.}
     The loss function for the $i^{th}$ training instance has the following form:
@@
 
-    \paragraph{Training the importance predictor.}
+    \paragraph{The importance predictor loss function.}
     There is a set of N training instances $R=\{(q_1, d_{1+}, d_{1-}), ..., (q_N, d_N+, d_N-)\}$,
@@
 
-    \paragraph{Training the gating controller.}
+    \paragraph{The gating controller loss function.}
     Consider a passage-query/summary parallel corpus \emph{C}, which contains passage \emph{p}, the
@@
 
-    \paragraph{Joint training.}
+    \paragraph{Joint loss function.}
     The importance predictor and gating controller are trained jointly because the supervisory
@@
 
-    \paragraph{Training}
+    \paragraph{Loss function.}
     During training, it is necessary to make a embeddings inner product for the relevant

### 2022-01-06 src/article.tex
Use dot at the end of the paragraph name.
Add training paragraph for SNRM.

@@
     \begin{equation}
+        \label{eqn:snrm:loss-function}
         \mathcal{L}(q_i,d_{i1},d_{i2},y_i) + \lambda\;L_1(\phi(q_i)||\psi(d_{i1})||\psi(d_{i2}))
@@
     \end{math}
-    ) and $||$ is concatenation sign.\\\par
+    ) and $||$ is concatenation sign.
+
+    \paragraph{Training.}
+    Weak supervision is used to train the model.
+
+    Adam is selected as the optimizer.
+    The learning rate is chosen from
+    ${\{5 \cdot 10^{-5}, 1 \cdot 10^{-4}, 5 \cdot 10^{-4}, 1 \cdot 10^{-3}, 5 \cdot 10^{-3}\}}$,
+    the batch size from \{32, 64, 128\}.
+    Ngram representation network has n equal to 5 and two or three hidden layers.
+    Hidden layer sizes and the output layer sizes are selected from \{100, 300, 500\} and
+    \{5k, 10k, 20k\}, respectively.
+    The hyper-parameter $\lambda$ (see Equation (\ref{eqn:snrm:loss-function})) is chosen from the
+    $[5 \cdot 10^{-9}, 1 \cdot 10^{-7}]$ interval.
+    The dropout rate is selected from \{0.6, 0.8, 1\}.\\\par
 
@@
 
-    \paragraph{Architecture}
+    \paragraph{Architecture.}
     \begin{figure}[h]
@@
 
-    \paragraph{Loss function}
+    \paragraph{Loss function.}
     The function presented in the Equation (\ref{eqn:sparterm:loss}) is supplemented with negative

### 2022-01-06 src/article.tex
Add training paragraph for SparTerm.

@@
 
+    \paragraph{Training.}
+    BERT (12-layer, 768-hidden, 12-heads, 110M parameters) is used for the Gating Controller
+    and Importance Predictor.
+    They are initialized with pre-trained models.
+    First, the Gating Controller is fine-tuned for 50k iterations where
+    $\lambda_1 = 1e - 3$ and $\lambda_2 = 1$.
+    Then Gating Controller parameters are fixed and SparTerm model is fine-tuned jointly for 100k
+    iterations.
+    Adam optimizer is selected with the learning rate $2 \cdot 10^{-5}$.\\\par
+
     Thus, SparTerm provides a function that is used to map the frequency-based and bag-of-words

### 2022-01-08 src/article.tex
Add training paragraph for SPLADE.

@@
 
+    \paragraph{Training.}
+    The models are initialized with the BERT-base pre-trained models.
+    Adam optimizer is selected with the learning rate $2 \cdot 10^{-5}$, linear scheduling,
+    a warmup of 6000 steps and a batch size of 124.
+    To reduce the contribution of the regularizer in the early stages of training, a planner for
+    $\lambda$ is used, which quadratically increases $\lambda$ at each training iteration, up to a
+    given step (50k), from which it is fixed for another 100k iterations.
+    The hyper-parameter $\lambda$ is chosen from the $[1 \cdot 10^{-4}, 1 \cdot 10^{-1}]$ interval.
+
     As a result, SPLADE is simple and effective model in information retrieval tasks.

### 2022-01-09 src/article.tex
Fix indentation.
Add training paragraph for RepBERT.

@@
         L = L_{rank} + L_{exp}
-    \end{equation}\\
+    \end{equation}
 
@@
     $\lambda_d$ are tunable hyper-parameters to control the sparsity of the query and the
-    document. \newline
+    document.
 
@@
     given step (50k), from which it is fixed for another 100k iterations.
-    The hyper-parameter $\lambda$ is chosen from the $[1 \cdot 10^{-4}, 1 \cdot 10^{-1}]$ interval.
+    The hyper-parameter $\lambda$ is chosen from the $[1 \cdot 10^{-4}, 1 \cdot 10^{-1}]$
+    interval.\newline
 
@@
     \emph{q} is query, \emph{m} is quantity of relevant documents and \emph{n - m} is quantity
-    of irrelevant documents\cite{RepBERT}.\\\par
+    of irrelevant documents\cite{RepBERT}.
+
+    \paragraph{Training.}
+    BERT base model is adapted.
+    It consists of 12 encoder layers with vector dimension of 768.
+    The model is fine-tuned for 350k iterations with batch size of 26 and gradient accumulation
+    steps of 2.
+    Adam optimizer is selected with the learning rate $3 \cdot 10^{-6}$,
+    $\lambda_1 = 0.9$ and $\lambda_2 = 0.999$, $L_2$ weight decay of 0.01, learning rate warmup
+    over the first 10k iterations, and linear decay of the learning rate.
+    Dropout probability is 0.1 on all layers.\newline
+
     The model has a simple architecture in order to improve work efficiency.
